---
title: "SEER ovarian cancer"
author: "Lizabeth Katsnelson"
date: "10/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE} 
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(caret)
library(cluster)
library(pROC)
library(randomForest)
library(tree)
library(caret)
library(gbm) 
library(PresenceAbsence)
library(glmnet)

```

### Clean Data
```{r}

# open data frame
data <-  read.csv(file="ovary_data_chemo.csv", header=TRUE, sep=",")

# change survival months to integer
data$Survival.months <- as.integer(data$Survival.months)

colnames(data)[which(names(data) == "Age.recode.with..1.year.olds")] <- "Agegroup"
colnames(data)[which(names(data) == "Race.recode..W..B..AI..API.")] <- "Race"
colnames(data)[which(names(data) == "Histology.recode...broad.groupings")] <- "Histology.Groups"
colnames(data)[which(names(data) == "Chemotherapy.recode..yes..no.unk.")] <- "Chemotherapy"
colnames(data)[which(names(data) == "Vital.status.recode..study.cutoff.used.")] <- "Vital.Status"



# change Race to character to add Hispanic labels - anyone who is labeled "white" or "unknown" for race and "hispanic" in origin recode column will be switched to "hispanic" in race column
data$Race <- as.character(data$Race)
data <- within(data, Race[Race == 'White' & Origin.recode.NHIA..Hispanic..Non.Hisp. == 'Spanish-Hispanic-Latino'] <- 'Spanish-Hispanic-Latino')
data <- within(data, Race[Race == 'Unknown' & Origin.recode.NHIA..Hispanic..Non.Hisp. == 'Spanish-Hispanic-Latino'] <- 'Spanish-Hispanic-Latino')
data$Race <- as.factor(data$Race)

# remove unwanted columns
data <- dplyr::select(data, -Patient.ID,
                      -SEER.registry,
                      -In.research.data,
                      -Origin.recode.NHIA..Hispanic..Non.Hisp.,
                      -Site.recode.ICD.O.3.WHO.2008,
                      -Behavior.recode.for.analysis,
                      -ICCC.site.recode.ICD.O.3.WHO.2008,
                      -Histologic.Type.ICD.O.3, 
                      -Behavior.code.ICD.O.3,
                      -ICD.O.3.Hist.behav..malignant,
                      -Summary.stage.2000..1998..,
                      -RX.Summ..Surg.Prim.Site..1998..,
                      -RX.Summ..Scope.Reg.LN.Sur..2003..,
                      -RX.Summ..Surg.Oth.Reg.Dis..2003..,
                      -Reason.no.cancer.directed.surgery,
                      -Scope.of.reg.lymph.nd.surg..1998.2002.,
                      -RX.Summ..Reg.LN.Examined..1998.2002.,
                      -Surgery.of.oth.reg.dis.sites..1998.2002.,
                      -Site.specific.surgery..1973.1997.varying.detail.by.year.and.site.,
                      -Radiation.to.Brain.or.CNS..1988.1997., 
                      -CS.tumor.size..2004..,
                      -COD.to.site.recode,
                      -SEER.cause.specific.death.classification,
                      -SEER.other.cause.of.death.classification,
                      -Survival.months.flag,
                      -COD.to.site.rec.KM,
                      -Type.of.follow.up.expected,
                      -Behavior.code.ICD.O.2,
                      -Histology.ICD.O.2, 
                      -Recode.ICD.O.2.to.9, 
                      -Recode.ICD.O.2.to.10, 
                      -Month.of.diagnosis.recode,
                      -State.county,
                      -ICD.O.3.Hist.behav)


# add 5 year survival variable column to set - yes or no
data$Five.Year.Survival <- ifelse(data$Survival.months >= 60, 'Yes', 
                                ifelse(data$Survival.months < 60, 'No', NA))
# change to factor
data$Five.Year.Survival <- as.factor(data$Five.Year.Survival)

# omit NA
data <- na.omit(data)

as.tibble(data) 
```

colnames(data)
Agegroup + Year.of.diagnosis + County + Race + State + Grade + Laterality + Diagnostic.Confirmation + 
Histology.Groups + SEER.historic.stage.A + Radiation.sequence.with.surgery + Radiation + Chemotherapy + 
Vital.Status + Year.of.birth + Month.of.diagnosis + Five.Year.Survival

Take out histology reports other than:
8010-8049: epithelial neoplasms, NOS
8050-8089: squamous cell neoplasms
8140-8389: adenomas and adenocarcinomas
8440-8499: cystic, mucinous and serous neoplasms

```{r}

data$Histology.Groups <- as.character(data$Histology.Groups)

data$Histology.Groups <- ifelse(data$Histology.Groups == '8010-8049: epithelial neoplasms, NOS' | data$Histology.Groups == '8050-8089: squamous cell neoplasms' | data$Histology.Groups == '8140-8389: adenomas and adenocarcinomas' | data$Histology.Groups == '8440-8499: cystic, mucinous and serous neoplasms', data$Histology.Groups, NA)
as.tibble(data$Histology.Groups)

#Histology.Groups[Histology.Groups == '8010-8049: epithelial neoplasms, NOS' | Histology.Groups == '8050-8089: squamous cell neoplasms' | Histology.Groups == '8140-8389: adenomas and adenocarcinomas' | Histology.Groups == '8440-8499: cystic, mucinous and serous neoplasms'] <- 'NA')

data$Histology.Groups <- as.factor(data$Histology.Groups)

as.tibble(data$Histology.Groups)

data %>%
  count(Histology.Groups)
```

```{r}
# omit NA rows (should get rid of 16,269 rows becuase we dont want most of the histology groups)
data <- na.omit(data)

as.tibble(data) 
```






```{r}

# get rid of survival months column (it is essentially the same as our classifier)
data <- dplyr::select(data, -Survival.months)

train_size <- floor(0.75 * nrow(data))
set.seed(12345)
train_pos <- sample(seq_len(nrow(data)), size = train_size)

train_classification <- data[train_pos, ]
test_classification <- data[-train_pos, ]

ctrl <- trainControl(method = "repeatedcv", repeats = 15,classProbs = T, savePredictions = T)
dim(train_classification)
dim(test_classification)
```


### Random Forest
```{r}
set.seed(343434)

RF_classification <- randomForest(Five.Year.Survival ~ ., data=train_classification, importance = TRUE, oob.times = 15, confusion = TRUE)

RF_classification
```


Visualize OOB error rate
```{r}
plot(RF_classification$err.rate[,1], type = "l", ylab = "Error rate", xlab = "Number of trees")
```

Visualize importance of features
```{r}
importance(RF_classification)
```

Predict using test set
```{r}
test_RF_classification <- predict(RF_classification , newdata = test_classification)

cm_rf <-  table(pred=test_RF_classification, true = test_classification$Five.Year.Survival)
cm_rf
```

```{r}
pcc(cm_rf, st.dev=TRUE)
```













### Classification Tree

```{r}
train_tree <- floor(0.2 * nrow(data))
set.seed(12345)
train_pos_tree <- sample(seq_len(nrow(data)), size = train_tree)

train_class_tree <- data[train_pos_tree, ]
test_class_tree <- data[-train_pos_tree, ]

dim(train_class_tree)
dim(test_class_tree)
```




```{r}
set.seed(12345)
classification_tree <- tree(Five.Year.Survival ~ Agegroup + Year.of.diagnosis + State + Grade + Laterality + Diagnostic.Confirmation + Histology.Groups + SEER.historic.stage.A + Chemotherapy + Month.of.diagnosis, data = train_class_tree, split = "gini")
plot(classification_tree)
text(classification_tree,cex=0.45)
```

```{r}
summary(classification_tree)
```


```{r}
classification_test <- predict(classification_tree, newdata = test_class_tree, type = "class")

confusionMatrix(classification_test, reference = test_class_tree$Five.Year.Survival)
```

```{r}
fit_classification_tree <- cv.tree(classification_tree,FUN=prune.misclass, K = 15)
```


```{r}
fit_classification_tree
```


```{r}
prune_classification_tree=prune.misclass(classification_tree, best = 2)

plot(prune_classification_tree)
text(prune_classification_tree,cex=0.45)
```


```{r}
classification_test_pruned <- predict(prune_classification_tree, newdata = test_classification, type = "class")

confusionMatrix(classification_test_pruned, reference = test_classification$Five.Year.Survival)
```












```{r, EDA}

data %>%
  count(Five.Year.Survival)

data %>%
  count(Histology.Groups)

ggplot(data = data) + 
  geom_histogram(mapping = aes(Histologic.Type.ICD.O.3))
ggplot(data = data) + 
  coord_flip() +
  geom_bar(mapping = aes(Laterality, fill = Vital.Status)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


ggplot(data = data) +
  geom_bar(mapping = aes(x = Agegroup, fill = Vital.Status)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


ggplot(data = data) +
  coord_flip() +
  geom_bar(mapping = aes(x = Race, fill = Vital.Status)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


ggplot(data = data) +
  coord_flip() +
  geom_bar(mapping = aes(x = Diagnostic.Confirmation, fill = Vital.Status)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


ggplot(data = data) +
  coord_flip() +
  theme(legend.position = 'none') +
  geom_bar(mapping = aes(x = Radiation.sequence.with.surgery, fill = Vital.Status)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

data %>%
  count(Radiation.sequence.with.surgery)

ggplot(data = data) +
    geom_bar(mapping = aes(x = Chemotherapy, fill = Vital.Status)) 

ggplot(data = data) +
  geom_bar(mapping = aes(x = Five.Year.Survival, fill = Chemotherapy))

```














